Towards a Neural Co-Processor Which Restores Movement After Stroke: Modeling a Proof-of-Concept

### 2021-09-27 paper/article.tex
Paper progress

\begin{abstract}
\textit{Objective} Brain co-processors\cite{rao.coproc} are devices which use artificial
intelligence (AI) and closed-loop neurostimulation to shape neural activity and to bridge
injured neural circuits for targeted repair and rehabilitation. The co-processor framework
offers a flexible approach to learning stimulation policies that optimize for
(a) specific regimes of neural activities, or (b) external task performance.
For example, it may seek to learn a complex stimulation policy, to aid a
stroke victim in reaching to grasp an object.
Through the use of artificial neural networks (ANNs) and deep learning, the
co-processor co-adapts with the neural circuit, allowing it to seek optimal
stimulation policies, and adapt them as the neural circuit changes. The results
presented here demonstrate a neural co-processor for the first time through
the use of a simulation, and explore some of the core algorithms that may allow them
to succeed. \textit{Approach} We provide the first proof-of-concept of a neural
co-processor, through the use of a simulated neural circuit, based on (ANNs).
The simulated circuit performs a reach-to-grasp task, based on visual
input which is designed to closely resemble a similar circuit in a primate brain
\cite{michaels.mrnn}. We simulate a variety of lesions by altering the model, and
then demonstrate a co-processor's ability to restore lost function through
``stimulation'' of that model. We further test the ability of a
co-processor to adapt its stimulation as the simulated brain
undergoes changes. \textit{Main results} Our simulated co-processor
successfully co-adapts with the neural circuit to accomplish the external reaching
task. The co-processor framework demonstrated here adapts to a variety of lesion
types, and to ongoing changes in the simulated brain. \textit{Significance} The proof-of-concept here
outlines a co-processor model, as well as our approach to training it, leading to
insights on how such a model may be developed for \textit{in vivo} use. We believe this
co-processor design will allow for learning complex stimulation policies that
help restore function to a stroke victim. This initial proof-of-concept lays
the foundation for a future demonstration \textit{in vivo}.
\end{abstract}

@@

\section{Introduction}

* Introduce the notion of a co-processor

* BCIs: a single direction
* Increasingly sophisticated ML techniques applied to decoding
** See 2019Raj
* Encoding models, stimulation models (cite)

@@

Need for more complex stimulation policies, and need for optimizing task performance.

Demonstration by way of simulation.

Simulation:
\begin{itemize}
	\item \cite{bernal.sim} Simulation of spiking neural network, learning stimulation regime
\end{itemize}

@@

* Overview of test bed
** Michaels model
** Lesion designs
** Stimulation design

>>>>

\begin{abstract}
\textit{Objective} Brain co-processors\cite{rao.coproc} are devices which use artificial
intelligence (AI) and closed-loop neurostimulation to shape neural activity and to bridge
injured neural circuits for targeted repair and rehabilitation. The co-processor framework
offers a flexible approach to learning stimulation policies that optimize for
(a) specific regimes of neural activities, or (b) external task performance.
For example, it may seek to learn a complex stimulation policy, to aid a
stroke victim in reaching to grasp an object.
Through the use of artificial neural networks (ANNs) and deep learning, the
co-processor co-adapts with the neural circuit, allowing it to seek optimal
stimulation policies, and adapt them as the neural circuit changes. The results
presented here demonstrate a neural co-processor for the first time, through
the use of a simulation, and explore some of the core algorithms that may allow them
to succeed. \textit{Approach} We provide the first proof-of-concept of a neural
co-processor, through the use of a simulated neural circuit, which  performs a
reach-to-grasp task, based on visual input. It is designed to closely resemble
a similar circuit in a primate brain \cite{michaels.mrnn}. We simulate a
variety of lesions by altering the model, and
then demonstrate a co-processor's ability to restore lost function through
``stimulation'' of that model. We further test the ability of a
co-processor to adapt its stimulation as the simulated brain
undergoes changes. \textit{Main results} Our simulated co-processor
successfully co-adapts with the neural circuit to accomplish the external reaching
task. The co-processor framework demonstrated here adapts to a variety of lesion
types, and to ongoing changes in the simulated brain. \textit{Significance} The proof-of-concept here
outlines a co-processor model, as well as our approach to training it, leading to
insights on how such a model may be developed for \textit{in vivo} use. We believe this
co-processor design will allow for learning complex stimulation policies that
help restore function to a stroke victim. This initial proof-of-concept lays
the foundation for a future demonstration \textit{in vivo}.
\end{abstract}

@@

\section{Introduction}
* Aided in part by application of advanced AI techniques, BCIs have made
advancements over the last several decads, allowing for decoded
brain signals to be used for control of a wide variety of virtual and physical
prostheses.
** Separately: advances in stimulation techniques and modeling have allowed us to probe neural circuit dynamics and learn to better drive neural circuits towards target dynamics (citations)
** Increasing interesting in combining decoding and encoding, for closed-loop control. BBCIs
** They have the potential to apply real-time, fine-grained control of neural circuits.
** Allows stimulation to be conditioned on a user's intention, sensory feedback, error signals, etc., to drive e.g. improvement in motor control after impairment due to TBI.

* Introduce the notion of a co-processor
** Use of AI and deep learning to determine a stimulation policy.
** Co-adapts policy to solve a task, together with the neural circuit
** Flexible model, allows for optimization of either brain activities or external task
** "These ``neural co-processors'' can be used to jointly optimize cost functions with the
   nervous system to achieve desired behaviors ranging from targeted neuro-rehabilitation
   to augmentation of brain function."
** Proposed approach here combines a learned stimulation model with a stimulation agent
** Learning allows for active adaptation to the neural circuit as it changes.

* BCIs: a single direction
* Increasingly sophisticated ML techniques applied to decoding
** See 2019Raj citations
* Encoding models, stimulation models (cite)

* Open loop control examples.  \cite{khanna.openloop}

* Learning stimulation models \cite{shanechi.stimmodel}

@@

* Need for more complex stimulation policies, and need for optimizing task performance.
** i.e. what if we don't have a specific pre-recorded activation we want to drive towards?
** And how do we find activations that drive an intended task solution?

Begin with a proof-of-concept in simulation.

@@

* Proof-of-concept by way of simulation.
** Simulation allows for rapid, cheap exploration of learning algorithms.
** Potentially helpful to look for a simulated neural circuit which has internal dynamics, i.e. information propagation, which is demonstrably similar to natural circuits.
** \cite{kao.sim} Likeness of RNNs to natural circuits re: dynamics
** Cite Michaels, and those upstream from him (Susillo?) re: RNNs
** \cite{bernal.sim} Simulation of spiking neural network, learning stimulation regime. Use of a biomimetic
  simulation to develop and evaluate a neural controller. Use of lesion model.

* Overview of test bed
** Michaels model \cite{michaels.mrnn}
** Lesion designs
*** Lesion examples, i.e. hand velocity more affected than shoulder velocity.
*** Disconnect modules vs lesion M1, and the effects
** Stimulation design
*** Spatio-temporal smoothing
